Research Notes

### 2019-02-05 presentations/meeting.tex
2/5/19 meeting?

\begin{frame}
    \frametitle{Linear}

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_0/fluxes.png}
        \caption{Linear simulation.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{New Plotting for Nonlinear}

    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/fluxes.png}
        \caption{Nonlinear simulation, new plotting methodology. Feedback?}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection?}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/p_074.png}
        \caption{Linear simulation with same viscosity as nonlinear. Green is
        viscosity-free envelope, orange analytical.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection (nl4)}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_074.png}
        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
        analytical.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection?}

    \begin{table}[t]
        \centering
        \begin{tabular}{l|l l }
            Name & RMS & Notes\\
            lin0 & 0.077 & Almost-no visc\\
            lin1 & 0.067 & Same visc as nl4\\
            nl1 (low) & 0.132 & half-res nl1, higher visc\\
            nl1 & 0.120 & $0.7\nu$\\
            nl2 & 0.213 & $0.4\nu$\\
            nl3 & 0.179 & $0.4\nu$ (double A)\\
            nl4 & 0.269 & $0.3\nu$\\
            nl5 & 0.207 & $0.3\nu$ (double $k_z$)\\
            nl6 & 0.246 & $0.2\nu$\\
            nl7 & 0.358 & $0.1\nu$
        \end{tabular}
        \caption{Table of RMS values.}
    \end{table}
\end{frame}

\begin{frame}
    \frametitle{Reflection (nl6)}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_6/p_074.png}
        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
        analytical.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection (nl7)}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_7/p_074.png}
        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
        analytical.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection (nl3)}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_3/p_058.png}
        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
        analytical.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Reflection (nl5)}

    Get exact solution, up to viscous dissipation. Plot $\delta u_z =
    \frac{u_{z,sim} - u_{z, anal}}{\abs*{u_{z, anal}}}$, ``fractional
    deviation'' between driving zone and critical layer. Compute $\mathrm{RMS} =
    \sqrt{\ev*{\delta u_z^2}}$.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_nl_5/p_074.png}
        \caption{Nonlinear simulation. Green is viscosity-free envelope, orange
        analytical.}
    \end{figure}
\end{frame}

>>>>

\begin{frame}
    \frametitle{Review}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item Linear: $\nu = 0$ has $S_{px}(z) = S_{px, 0}$ great
                    prediction. Fit $\vec{u}$ for $\nu \neq 0$ as well, though
                    not $S_{px}(z)$ yet (\textbf{Updated!}).

                \item In nonlinear, review: expect deposited flux $\Delta S_{px}
                    = S_{px, 0}$ incident/generated flux.

                \item Instead, observe $\Delta S_{px} < S_{px, 0}$, and moreover
                    $S_{px}$ far from critical layer changing over time!
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item Last week: defined some
                    \begin{equation}
                        \delta u_z = \frac{u_z - u_{z, 0}
                            - \ev*{u_z}_x}{\abs*{u_{z, 0}}},
                    \end{equation}
                    deviation (new: subtracting mean flow too), computed RMS\@.
                    Found higher RMS for lower $\nu$.

                \item Could it be $\ev*{\dots}_x$ vs $\ev*{\dots}_t$?

                \item Objective: understand ``reflection''?
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Overview}

    \begin{itemize}
        \item Understand exactly how much flux is being absorbed at the critical
            layer:
            \begin{itemize}
                \item Can the decreased flux absorption be explained by
                    viscous dissipation before the critical alyer alone?

                \item Compare predictions of $\pd{z_c}{t}$ to the observed
                    $\Delta S_{px}$ at the critical layer.
            \end{itemize}

        \item Understand nature of ``reflection'', why cannot see in certain
            simulations.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Linear}

    Had great $u_z$ fits including viscosity (error $\sim 3-5\%$), but did not
    show great $S_{px}(z; \nu)$ fits last week. Here they are!
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.7\textwidth]{../sims/2d_3_final/snapshots_lin_1/fluxes.png}
        \caption{Fits linear $S_{px}(z; \nu)$.}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Mean Flow and Flux}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fluxes.png}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item Last week: Evolution of $\ev*{u_x}_x(t),
                    \ev*{S_{px}}_x(t)$ at different times. Note flux seems to
                    decrease over time throughout domain.

                \item Bottom panel, solid lines are time slices, dotted lines
                    indicate averaging over $\sim \frac{2\pi}{\omega}$ one
                    period (6 times), thick line is linear viscous prediction.

                \item Shaded grey indicates one vertical wavelength.
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Mean Flow and Flux}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fluxes.png}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item More damped for comparison ($\nu \to 7\nu/3$).

                \item Note that $S_{px}$ does not change much over time, general
                    trend matches viscous linear decrease.
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Absorbed Flux and Front Propagation}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/front.png}
            \end{figure}
            \begin{itemize}
                \item Three different $\Delta S_{px}$:
                    \begin{itemize}
                        \item LinVisc extrapolate $S_{px, 0}$.

                        \item $\Delta S_{px; S}$ ($z$-avg).

                        \item $\Delta S_{px; U}$ ($z$-avg too).
                    \end{itemize}
            \end{itemize}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}

                \item \textbf{FIXED} $z_c(t)$ for different incident fluxes,
                    even more models:
                    \begin{itemize}
                        \item Solid lines are three ways of extracting from
                            data.

                        \item Dotted lines use $\pd{z_c}{t}(\Delta S_{px})$
                            model for three values of $S_{px}$:
                            \begin{itemize}
                                \item Average $\Delta S_{px}$.
                                \item $\nu$-extrapolated $S_{px, 0}$.
                                \item Full $S_{px, 0}$.
                            \end{itemize}
                    \end{itemize}
                    Takeaway: strongly inconsistent with $S_{px, 0}$ being
                    completly absorbed! In fact, most are $\sim 0.5S_{px}$ being
                    absorbed.
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Absorbed Flux and Front Propagation}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/front.png}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item Overdamped for comparison.

                \item Absorbed flux is in good agreement w/ linear viscous
                    extrapolation of $S_{px, 0}$, seems to imply full
                    absorption.
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Residual Flow Properties}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/fft.png}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{itemize}
                \item FT of $\delta \vec{u}$ slightly below the critical layer.
                    Red line denotes where $2\nu k_x^2 = \omega$, ``viscous
                    wavenumber''?

                \item Note that we've chosen $\nu k_x^2 \lesssim u_xk_x \sim
                    \frac{\omega}{k_{0x}}k_x$, our previous viscous wavenumber
                    already.
            \end{itemize}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Results}
    \framesubtitle{Residual Flow Properties}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/fft.png}
                \caption{Overdamped for comparison, not power law!}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_lin_2/fft.png}
                \caption{Linear for comparison; taken just below damping zone.}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}
    \frametitle{Useful Plots}
    \framesubtitle{Residuals}

    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_4/p_009.png}
                \caption{Residuals for not overdamped. }
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}[t]
                \centering
                \includegraphics[width=\textwidth]{../sims/2d_3_final/snapshots_nl_1/p_009.png}
                \caption{Residuals for overdamped.}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

### 2019-02-05 writeups/notes.tex
2/5/19 meeting?

>>>>

\section{Estimating Various Other Scales}

\subsection{Viscous Damping Length}

This is a rather hacky derivation. Consider that in the fluid equations, if we
assume $\frac{1}{k_zH} \ll 1$, then viscosity basically just enters as a bunch
of $\nu k^2$ terms. We can absorb this into the frequency $\partial_t + \nu k^2
\sim i\omega_0\p*{1 - \frac{i\nu k^2}{\omega_0}}$. Plugging into the dispersion
relation then shows that ($k_0 = \abs*{\vec{k}_0} = \sqrt{k_{x0}^2 + k_{z0}^2}$)
\begin{align*}
    \omega^2 &\approx \frac{N^2k_{x0}^2}{k^2}\p*{1 - \frac{2i\nu k^2}{\omega_0}},\\
        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 k_{z0}^2\p*{1 + \frac{2i\nu k^4}{\omega
            k_{z0}^2}}},\\
        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 + \s*{k_{0z}
            \p*{1 + \frac{i\nu k_0^4}{\omega k_{0z}^2}}}},\\
        &\approx \frac{N^2k_{x0}^2}{k_{x0}^2 + \s*{k_{0z}
            \p*{1 + \frac{i\nu k_0^5}{Nk_{x0} k_{0z}^2}}}},\\
    e^{ik_zz} &\approx \exp\s*{ik_{z0}z - \frac{\nu k_0^5}{Nk_{x0}k_{z0}}z}.
\end{align*}
Thus, in the small dissipation limit we can treat the artificial viscosity the
same as a vertical damping length.
